Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-11-01 manuscript/manuscript.Rmd
WIP

```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 \cdot 10^4$.")

@@

Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions. Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option to simulate realistic feature distributions. This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:

@@

The optimal weighting function parameters for weighted MLE are shown in Tables **insert** and **insert** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.

@@

Figure **insert** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters. In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation. For both transformations, weighting based on empirical probabilities yields the most consistent $\lambda$ parameter estimates in the presence of outliers. Weighting based on residual errors tended to perform least well, and were generally similar to (Yeo-Johnson) or worse than (Box-Cox) non-robust transformations.

@@

>>>>

```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 \\cdot 10^4$.")

@@

Modifications intended to make power transformations invariant to shifts and methods to improve their robustness against outliers need to be assessed using data drawn from a range of different distributions. Since the power transformations are intended for use with unimodal distributions, the generalised normal distribution [@Subbotin1923-qk; @Nadarajah2005-xe] is a suitable option for simulating realistic feature distributions. This distribution has the following probability density function $f_{\beta}$ for a value $x \in \mathbb{R}$:

@@

The optimal weighting function parameters for weighted MLE are shown in Tables **INSERT** and **INSERT** for robust, shift-sensitive Box-Cox and Yeo-Johnson transformations, respectively.

@@

Figure **INSERT** shows the distribution of errors $\left| \hat{\lambda}^r - \hat{\lambda}_0 \right|$ for non-robust transformations, as well as robust transformations using the optimal weighting function parameters. In the presence of outliers on either side of the distribution, the Yeo-Johnson transformation yields smaller errors than the Box-Cox transformation. For both transformations, weighting based on empirical probabilities yields the most consistent $\lambda$ parameter estimates in the presence of outliers. Weighting based on residual errors tended to perform least well, and were generally similar to (Yeo-Johnson) or worse than (Box-Cox) non-robust transformations.

@@

The Top Gear dataset contains data on 297 cars that appeared on the BBC television show *Top Gear*. Fuel consumption contains outliers.

```{r top-gear, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Insert caption here"
)

.plot_top_gear(
  plot_theme = plot_theme
)
